---
title: "PANAS by country preliminary analysis"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
```


```{r libraries}
library(dplyr)
library(tidyverse)
library(here)
library(DataCombine)
```

Specify paths for each of the files and load them
```{r loaddata, include=TRUE}
dict_path <- here("data", "capstone20_dictionaries.csv")  
adjcount_path <- here("data", "dictionary_word_matrix_adj.csv")  
countries_path <- here("data", "countries-languages.csv")
mh_path <- here("data", "country-mentalhealth.csv")

# List of dictionaries in dataset
dict <- read_csv(dict_path)  

# List of countries and the languages they speak
lang_countries <- read_csv(countries_path) 

# Mental health disorder prevalance by country
mh <- read_csv(mh_path) 
```

Positive and negative affect words from PANAS X
```{r loadwords, include=TRUE}
positive_words = c("active","alert","attentive","enthusiastic","excited","interested","proud","happy", "joyful","delighted","cheerful","lively","energetic","strong","confident","bold","daring","fearless")
negative_words <- c("afraid","scared","nervous","jittery","irritable","hostile","guilty","ashamed","upset", "distressed","frightened","shaky","angry","disgusted","sad","lonely")
```



Calculate mean frequencies for positive and negative affect
```{r calculatefrequencies, include=TRUE}
# load adj matrix and sum across columns to compute the total size of each dictionary
adjcount <- read_csv(adjcount_path) %>%  
  mutate(dictsize = select(., abbreviated:zealous) %>% apply(1, sum, na.rm=TRUE))

word_count <- adjcount %>% 
  select(id, positive_words, negative_words, dictsize) %>% mutate_all(~replace_na(.x, 0))

# we'll express frequencies in terms of counts per 10000 tokens
unit <- 10000

# 
word_frequencies <- dict %>% 
  inner_join(word_count, by="id") %>%
  filter(dictsize >= 5000)  # only include dictionaries with more than 5000 tokens

# generate frequencies
for (word in c(positive_words, negative_words)){
  word_frequencies <- word_frequencies %>% mutate(!!sym(word) :=  unit * !!sym(word)/dictsize)
}

# calculate mean frequencies
word_frequencies$mean_positive<-rowSums(word_frequencies[,positive_words])
word_frequencies$mean_negative<-rowSums(word_frequencies[,negative_words])

# take the mean of the frequencies for each dictionary in order to merge into a single language
panas_frequencies <- aggregate(word_frequencies, by = list(language = word_frequencies$langname), FUN = mean)

#select relevant columns
panas_frequencies <- panas_frequencies %>% select(language, mean_positive, mean_negative)
```


Replace Language Names to match CIA dataset to Dictionary Dataset
```{r replacelanguagenames, include=TRUE}
replaces <- data.frame(from = c("Albanian", "Bosnian", "Khmer", "Croatian", "Serbian", "Arabic", "Mandarin", "Persian", "Mongolian", "Greek", "Hebrew", "Armenian", "Malay", "Kyrgyz"), to = c("Northern Tosk Albanian", "Serbian-Croatian-Bosnian", "Central Khmer", "Serbian-Croatian-Bosnian", "Serbian-Croatian-Bosnian", "Standard Arabic", "Mandarin Chinese", "Western Farsi", "Halh Mongolian", "Modern Greek", "Modern Hebrew", "Eastern Armenian", "Standard Malay", "Kirghiz"))
lang_countries <- FindReplace(data = as.data.frame(lang_countries), Var = "Language", replaceData = replaces,
                     from = "from", to = "to", exact = TRUE)

```


Merge IMHE data with CIA data to obtain disorder prevelance by countries and languages spoken
```{r createtemp, include=TRUE}
# Obtain Depressive Disorder Prevalence by Country
depression_prevalence <- mh %>%   select(location, val, cause)

temp1 <- merge(x= depression_prevalence, y = lang_countries, by.x = "location", by.y = "Country")

```


Merge temporary dataset with panas frequencies
```{r createdataframe, include=TRUE}
d <- merge(x= temp1, y = panas_frequencies, by.x = "Language", by.y = "language")

# weigh each country by the percentage of people speaking the particular language
d <- d %>% group_by(location) %>% mutate(mean_positive = weighted.mean(mean_positive, Percentage/100)) %>% mutate(mean_negative = weighted.mean(mean_negative, Percentage/100)) %>% mutate(Percentage = sum(Percentage)/4) %>% subset(select = -c(Language)) %>% distinct()

# remove countries with percentage < 80
d <- d %>% subset(Percentage >= 80) %>% subset(select = -c(Percentage))

# convert val to percentage
d <- d %>% mutate(val = val*100)


# add ratio column and total column
d <- d %>% mutate(ratio = mean_negative / mean_positive)
d <- d %>% mutate(total = mean_negative + mean_positive)

#pivot longer
d <- d %>% pivot_wider(names_from = "cause", values_from="val")

# rename columns
d <- rename(d, c("depressive_disorders"="Depressive disorders", "dysthymia"="Dysthymia",  "major_depressive_disorder"="Major depressive disorder",  "anxiety_disorders" = "Anxiety disorders"))
```

## Exploratory Plot
Draw graph for depressive disorders (all three disorders combined)
```{r plotgraphs, include=TRUE}

plotd <- d %>% 
  pivot_longer(cols=c(mean_positive, mean_negative, ratio, total), names_to="affect_words", values_to="propn") %>% 
  mutate(affect_words = recode(affect_words, mean_positive="positive words", mean_negative="negative words", ratio="ratio of negative to positive", total="total of negative and positive")) 


ggplot(plotd, aes(x=depressive_disorders, y=propn)) +
  geom_point() +
  facet_wrap(~affect_words, scales = "free") +
  geom_smooth(method=lm) +
  labs(x="depressive disorder prevalence/%", y="counts per 1000, ratio")

```



Top Right:
This plot suggests that mean frequency of positive emotion words is associated with higher prevalence of depressive disorders (each datapoint is a country). The x-axis represents depressive disorder prevalence in each country. The y-axis represents the mean frequency of positive words in language(s) that are spoken by >= 80% of the population in that country. 


Top left:
This graph plots the relationship between count of negative PANAS words on the y-axis  and the countries’ depressive disorder prevalence (percentage of total population) on the x-axis. As we see an increase in negative PANAS word counts, an increase in depressive disorder prevalence is also observed. Overall, this plot indicates a greater amount of negative PANAS words may be related to an increased prevalence of depressive disorder. 



Bottom left: 
This plot shows the correlation between the ratio of negative to positive words present in a language (y-axis) and the depressive disorder prevalence of the country in which that language is predominantly spoken (x-axis). It suggests a negligible relationship between the two variables, with an increase in the depressive disorder prevalence being associated with a negligible change in the ratio of negative to positive words.


Bottom Right: This plot shows the correlation between the total count of both positive and negative words present in a language (y-axis) and the depressive disorder prevalence of the country in which that language is predominantly spoken (x-axis). It suggests a positive relationship between the two variables, with an increase in the depressive disorder prevalence of a country being associated with an increase in the total count of positive and negative words in the language predominantly spoken in that country.

## Further Plots
Draw graphs for other depressive disorders
```{r plotgraphs2, include=TRUE}
ggplot(plotd, aes(x=dysthymia, y=propn)) +
  geom_point() +
  facet_wrap(~affect_words, scales = "free") +
  geom_smooth(method=lm) +
  labs(x="dysthymia prevalence/%", y="counts per 1000, ratio")


ggplot(plotd, aes(x=major_depressive_disorder, y=propn)) +
  geom_point() +
  facet_wrap(~affect_words, scales = "free") +
  geom_smooth(method=lm) +
  labs(x="major depressive disorder prevalence/%", y="counts per 1000, ratio")


ggplot(plotd, aes(x=anxiety_disorders, y=propn)) +
  geom_point() +
  facet_wrap(~affect_words, scales = "free") +
  geom_smooth(method=lm) +
  labs(x="anxiety disorder prevalence/%", y="counts per 1000, ratio")
```


It was found that similar trends were seen in the graphs for anxiety disorders and major depressive disorders, but a different trend was seen for dysthymia.


## Control Independant Variable
Select 36 random words and perform same method to compare with PANAS words
```{r calculatecontrol, include=TRUE}
random_words <- colnames(sample(select(adjcount, abbreviated:zealous), size=36))

control_word_count <- adjcount %>% 
  select(id, random_words, dictsize) %>% mutate_all(~replace_na(.x, 0))

# we'll express frequencies in terms of counts per 10000 tokens
unit <- 10000

# 
control_word_frequencies <- dict %>% 
  inner_join(control_word_count, by="id") %>%
  filter(dictsize >= 5000)  # only include dictionaries with more than 5000 tokens

# generate frequencies
for (word in random_words){
  control_word_frequencies <- control_word_frequencies %>% mutate(!!sym(word) :=  unit * !!sym(word)/dictsize)
}

# calculate mean frequencies
control_word_frequencies$mean_words<-rowSums(control_word_frequencies[,random_words])

# take the mean of the frequencies for each dictionary in order to merge into a single language
control_frequencies <- aggregate(control_word_frequencies, by = list(language = control_word_frequencies$langname), FUN = mean)

#select relevant columns
control_frequencies <- control_frequencies %>% select(language, mean_words)

d_control <- merge(x= temp1, y = control_frequencies, by.x = "Language", by.y = "language")

# weigh each country by the percentage of people speaking the particular language
d_control <- d_control %>% group_by(location) %>% mutate(mean_words = weighted.mean(mean_words, Percentage/100)) %>% mutate(Percentage = sum(Percentage)/4) %>% subset(select = -c(Language)) %>% distinct()

# remove countries with percentage < 80
d_control <- d_control %>% subset(Percentage >= 80) %>% subset(select = -c(Percentage))

# convert val to percentage
d_control <- d_control %>% mutate(val = val*100)

#pivot longer
d_control <- d_control %>% pivot_wider(names_from = "cause", values_from="val")

# rename columns
d_control <- rename(d_control, c("depressive_disorders"="Depressive disorders", "dysthymia"="Dysthymia",  "major_depressive_disorder"="Major depressive disorder",  "anxiety_disorders" = "Anxiety disorders"))

plotd_control <- d_control %>% 
  pivot_longer(cols=c(depressive_disorders, major_depressive_disorder, anxiety_disorders, dysthymia), names_to="disorders", values_to="prevalence") 

print(random_words)

ggplot(plotd_control, aes(x=prevalence, y=mean_words)) +
  geom_point() +
  facet_wrap(~disorders, scales = "free") +
  geom_smooth(method=lm) +
  labs(x="depressive disorder prevalence/%", y="counts per 1000, ratio")

```


Further tests can be done by calculating the chance of a significant result using 36 random words from the dictionary dataset over thousands of sets of random words